A Standardized Effect Size for Evaluating the Strength of Phylogenetic Signal, and Why Lambda is not Appropriate

### 2020-04-30 Manuscript/Manuscript-4.Rmd
vsn 4

@@

Evolutionary biologists commonly seek to describe the relative strength of phylogenetic signal in phenotypic traits, to determine the extent to which shared evolutionary history has influenced trait covariation among taxa. This is often accomplished by interpreting empirical estimates of $\lambda$; with smaller values signifying 'weak' phylogenetic signal, while larger values are interpreted as 'strong' phylogenetic signal [e.g., @DeMeester2019; @Pintanel2019;@Su2019]. Other approaches for interpreting $\lambda$ are more statistical. For instance, some have evaluated whether the observed $\lambda$ differs from some expected value through the use of confidence intervals [@Vandelook2019] or by performing likelihood ratio tests that compare the observed model fit to that obtained when $\lambda=0$ or $\lambda=1$ [@Freckleton_et_al2002; @Cooper2010; @Bose2019]. Additionally, qualitative comparisons of $\lambda$ estimates obtained from multiple phenotypic traits have been used to infer whether the strength of phylogenetic signal is greater in one trait as compared to another [e.g., @Liu2019; @Bai2019]. Indeed, statements regarding the strength of phylogenetic signal based on $\lambda$ are rather common in the evolutionary literature. For instance, of the 204 papers published in 2019 that estimated and reported Pagel's $\lambda$ (found in Google.scholar), 40% interpreted the strength of phylogenetic signal for at least one phenotypic trait. Additionally, because nearly half of the 1,572 $\lambda$ values reported were near the limits of the parameter (Figure 1), this percentage is even higher, as biological interpretation of phylogenetic signal at the limits of $\lambda$ are known. \hfill\break

@@

It seems intuitive to interpret the strength of phylogenetic signal based on the value of $\lambda$, as $\lambda$ is a parameter on a bounded scale ($0\to1$) for which interpretation of its extremal points are understood. Specifically, $\lambda=0$ represents no phylogenetic signal, while $\lambda=1$ is phylogenetic signal as expected under Brownian motion. However, equating values of $\lambda$ directly to the strength of phylogenetic signal presumes two important statistical properties that have not been fully explored. First, it presumes that values of $\lambda$ can be precisely estimated, as biological inferences regarding the strength of phylogenetic signal depend on high accuracy in its estimation. Therefore, understanding the precision in estimating $\lambda$ is paramount. One study [@Boettiger_et_al2012] found that estimates of Pagel's $\lambda$ displayed less variation (i.e., greater precision) when data were simulated on a large phylogeny ($N=281$) as compared to a small one ($N=13$). From this observation it was concluded that insufficient data (i.e., the number of species) was the underlying cause of the increased variation across parameter estimates [@Boettiger_et_al2012]. Indeed, such a pattern is common with statistical estimators, as summary statistics and parameters are often more precise at greater sample sizes [@Cohen1988]. However, this conclusion also assumes that the precision of $\lambda$ remains constant across its range ($\lambda = 0 \to 1$); an assumption that to date, has not been verified. Thus, despite widespread use of Pagel's [-@Pagel1999] $\lambda$ in macroevolutionary studies, at present, we still lack a general understanding of the precision with which $\lambda$ can estimate levels of phylogenetic signal in phenotypic datasets. \hfill\break

Second, while estimates of $\lambda$ are within a bounded scale ($0\to1$), this does not *de-facto* imply that the estimated values of this parameter correspond to the actual strength of the underlying input signal in the data. For this to be the case, $\lambda$ must be a statistical effect size. Effect sizes are a measure the magnitude of a statistical effect in data, represented on a common scale [@Glass1976; @Cohen1988]. Effect sizes have widespread use in many areas of the quantiative sciences, as they represent measures that may be readily summarized across datasets as in meta-analysis [@HedgesOlkin1985; @Glass1976; @Arnqvist1995], or compared among datasets [e.g., @AdamsCollyer2016; @AdamsCollyer2019b]. Unfortunatley, not all model parameters and test statistics are effect sizes, and thus many summary measures must first be converted to standardized units (i.e., an effect size) for meaningful comparison [see @Rosenthal1994]. As a consequence, it follows that only if $\lambda$ is a statistical effect size can comparisons of estimates across datasets be interpretable. For the case of $\lambda$, this has not yet been explored.  \hfill\break

In this study, we evaluate the precision of Pagel's $\lambda$ for estimating known levels of phylogenetic signal in phenotypic data. We use computer simulations with differing numbers of species, differently shaped phylogenies, and differing input levels of phylogenetic signal, to explore the degree to which $\lambda$ correctly identifies known levels of phylogenetic signal, and under what circumstances. We find that estimates of $\lambda$ vary widely for a given input value of phylogenetic signal, and that the precision in estimating $\lambda$ is not constant across its range. Rather, there is decreased precision when input levels of phylogenetic signal are of intermediate strength. Additionally, the same estimated values of $\lambda$ may be obtained from datasets containing vastly different input levels of phylogenetic signal. Thus, $\lambda$ is not a reliable estimate of the strength of phylogenetic signal in phenotypic data. We then describe a standardized effect size for measuring the strength of phylogenetic signal in phenotypic datasets, and apply the concept to two common measures of phylogenetic signal: $\lambda$ and $Kappa$. Through simulations across a wide range of conditions, we find that the precision of effect sizes based on $\lambda$ ($Z_{\lambda}$) are less reliable than that those based on $Kappa$ ($Z_K$), implying that $Z_K$ is a more robust effect size measure. We also propose a two-sample test statistic that may be used to compare the strength of phylogenetic signal among datasets, and provide an empirical example to demonstrate its use. We conclude that estimates of phylogenetic signal using Pagel's $\lambda$ are often inaccurate, and thus interpreting strength of phylogenetic signal in phenotypic datasets based on this measure is compromised. By contrast, effect sizes obtained from $Kappa$ hold promise for characterizing phylogenetic signal, and for comparing the strength of phylogenetic signal across datasets.

@@

We conducted a series of computer simulations to evaluate the precision of Pagel's $\lambda$. Our primary simulations were based on pure-birth phylogenies; however, we also evaluated patterns on both balanced and pectinate trees to determine whether tree shape affected our findings (see Supporting Information). First we generated 50 pure-birth phylogenies at each of six different tree sizes, ranging from 32 to 1024 taxa ($n=2^5 - 2^{10}$). Next, we rescaled the simulated phylogenies by multiplying the internal branches by $\lambda_{in}$, using 21 intervals of 0.05 units across its range ($\lambda_{in} = 0.0 \to 1.0$), resulting in 1050 scaled phylogenies at each level of species richness ($n$). Continuous traits were then simulated on each phylogeny under a Brownian motion model of evolution to obtain datasets with differing levels of phylogenetic signal, that ranged from no phylogenetic signal (when $\lambda_{in} =0$), to phylogenetic signal corresponding reflecting Brownian motion (when $\lambda_{in} =1$). For each dataset we then estimated phylogenetic signal ($\lambda_{est}$), and calculated the precision of $\lambda$ using the variance ($\sigma^2_\lambda$) across datasets at each input level of phylogenetic signal and level of species richness. \hfill\break

We also evaluated the precision of $\lambda$ when estimated in PGLS regression and ANOVA (i.e., $Y\sim{X}$). Here, an independent variable $X$ was simulated on each phylogeny under a Brownian motion model of evolution (for PGLS regression). For phylogenetic ANOVA, random groups ($X$) were obtained by simulating a discrete (binary) character on each phylogeny. Next, the dependent variable was simulated in such a manner as to contain a known relationship with $X$ plus random error containing phylogenetic signal. This was accomplished as: $Y=\beta{X}+\epsilon$. Here, the association between $Y$ and $X$  was modeled using a range of values: $\beta=(0.0,0.25, 0.5, 0.75,1.0)$, and the residual error was modeled to contain phylogenetic signal simulated under a Brownian motion model of evolution: $\epsilon=\mathcal{N}(\mu=0,\sigma=\mathbf{C})$: [see @Revell2010 for a similar simulation design]. The fit of the phylogenetic regression was estimated using maximum likelihood, and parameter estimates ($\beta_{est}$ and $\lambda_{est}$) were obtained. Precision estimates ($\sigma^2_\lambda$) at each input level of phylogenetic signal and level of species richness were then observed. \hfill\break

@@

*Results.* We found that the precision of $\lambda_{est}$ varied widely across simulation conditions. Predictably, precision improved as the number of species increased (Figure 2). This confirmed earlier findings of Boettiger et al. [-@Boettiger_et_al2012], and adhered to parametric statistical theory. However, in many cases the set of $\lambda_{est}$ spanned nearly the entire range of possible values (e.g., $n=32$; $\lambda_{in}=0.5$: $\lambda_{est}= 0.0\to 0.985$), revealing that estimates of $\lambda$ were not a reliable indicator of input phylogenetic signal. Importantly, the precision of $\lambda_{est}$ was not uniform across all levels of phylogenetic signal. The worst precision was observed at intermediate levels of phylogenetic signal ($\lambda_{in}\approx0.5$), while precision improved as input levels approached the extremes of $\lambda$'s range (i.e., $\lambda_{in}\rightarrow0$ & $\lambda_{in}\rightarrow1$). Thus, estimates of $\lambda$ were least reflective of the true input signal at intermediate values. Additionally, even at large levels of species richness, we found that the range of $\lambda_{est}$ still encompassed a substantial portion of possible values (e.g., $n=512$; $\lambda_{in}=0.5$: $\lambda_{est} = 0.32\to 0.68$). Likewise, the same $\lambda_{est}$ could be obtained from datasets containing vastly different input levels of phylogenetic signal (e.g., $n=512$; $\lambda_{est} = 0.5$; $\lambda_{in} = 0.25 \to 0.65$). These findings were particularly unsettling when considered in light of our literature survey. Over one quarter of the $\lambda$ estimates obtained in empirical studies (421 of 1,572) were between $\lambda=0.25$ and $\lambda=0.75$ (Figure 1). This range reflected the region that our simulations identified as being the least reliable in terms of accurately characterizing levels of phylogenetic signal, yet 30% of these empirical estimates were explicitly interpreted in terms of the strength of phylogenetic signal that they represented (i.e., weak, intermediate, strong phylogenetic signal). \hfill\break

Finally, when $\lambda$ was co-estimated with regression parameters in PGLS regression, the results of our simulations were quite similar. Here, regression parameters ($\beta$) were accurately estimated, confirming earlier findings of Revell [-@Revell2010] (see Supporting Information). However, estimates of phylogenetic signal ($\lambda$) were less precise (Figure 3), and the spread of $\lambda_{est}$ was similar to that observed when $\lambda$ was estimated for only the dependent variable, as in Figure 2. Taken together, these findings reveal that $\lambda_{est}$ does not precisely characterize observed levels of phylogenetic signal in phenotypic datasets, and that biological interpretations of the strength of phylogenetic signal based on $\lambda$ may be highly inaccurate. \hfill\break

@@

Adams and Collyer [-@AdamsCollyer2019] suggested that the strength of phylogenetic signal could be represented as an effect size, based on the $Kappa$ statistic and its empirical sampling distribution from permutation. Here we formalize that suggestion, resulting in an effect size of:

\begin{align}
    Z_{K}=\frac{K_{obs}-\hat\mu_K}{\hat\sigma_K}
\end{align}

where $K_{obs}$ is the observed phylogenetic signal, and $\hat\mu_K$ and $\hat\sigma_K$ are the mean and standard deviation of the empirical sampling distribution of $Kappa$ obtained via permutation. Similarly, an effect size based on $\lambda$ could be envisioned as:

@@

In this case, $\lambda_{obs}$ and $\hat\sigma_\lambda$ are empirically derived using maximum likelihood. Note that under the null hypothesis, no phylogenetic signal is expected [@Freckleton_et_al2002], and thus $E(\lambda)=0$ under this condition. \hfill\break

To evaluate the utility of $Z_K$ and $Z_\lambda$ we calculated both effect sizes for the simulated datasets generated above, and summarized the precision of each using its variance ($\sigma^2_{Z_K}$ and $\sigma^2_{Z_\lambda}$). Results are found in Figure 4 (additional results are found in the Supporting Information). Here two things are evident. First, estimates of $Z_K$ track the input phylogenetic signal in a more linear fashion than do estimates of $Z_\lambda$ (Figure 4A,B). Thus, actual changes in the strength of phylogenetic signal are reflected more evenly in the corresponding values of the effect size $Z_K$. Second, the precision of $Z_K$ is considerably more stable as compared with $Z_\lambda$. This may be seen by calculating the coefficients of variation for the set of precision estimates (i.e., $\sigma^2_{Z_K}$ and $\sigma^2_{Z_\lambda}$) across input levels of phylogenetic signal.  Here coefficients of variation in the precision of $Z_K$ were an order of magnitude smaller for than for $Z_\lambda$ (Figure 4C,D). This implied that estimates of the strength of phylogenetic signal were more reliable and robust when using $Z_K$ as compared with $Z_\lambda$. \hfill\break

@@

Once the magnitude of phylogenetic signal is characterized using $Z_K$, one may wish to compare such measures across datasets, to determine whether the strength of phylogenetic signal is significantly greater in one phenotypic trait as compared with another. As with other effect sizes derived from permutation distributions [e.g., @AdamsCollyer2016; @AdamsCollyer2019b], a two-sample test statistic may be found as:

@@

One important advantage of the approach advocated here is that the resulting effect sizes ($Z_K$) are dimensionless, as the units of measurement cancel out during the calculation of $Z$ [@SokalRohlf2012]. As a consequence, the effect sizes represent the strength of phylogenetic signal on a common and comparable scale, measured in standard deviation units, regardless of the initial units and scale of the original phenotypic variables under investigation. This means that conceivably, the strength of phylogenetic signal may be compared across datsets for phenotypic traits in different units and scale (e.g., behavioral and morphological traits), so long as they are continuous variables. Note that this is directly related to the common practice of converting data to their standard normal deviates prior to their use in multivariate analysis, if the original variables are expressed in different units or scales [see @Everitt2011; @LegendreLegendre2012; @AdamsCollyer2019].

@@

# Conclusions and Implications

1: summary paragraph

2: expand on Lambda..
lambda innacurate, not precise, level of precision varies with input physig (worse in mid-range). NEW RESULT.  We are first to show this.  NOTE: pattern is obvious with reflection. Since it is a 'bounded' parameter estimation should be best at the extremes... (state this?).. hmm.

Patterns worse with PGLS, though beta still estimated properly. Conclusion, lambda not overly useful.

3: By contrast, effect size Z-K useful, equally precise across range of values. Can be used to characterize the strength of physignal, and because robust to input levels, etc. may be used to compare across datasets.

Somewhere, recognize that this is somewhat 'backwards' from prior recommendations where Kappa had somewhat lower performance in terms of type I and type II error (which?? I forget). However, recall that those studies did not examine the precision of the estimates. Nor was Z-k included, because it was not yet invented. So Use of Z-k should make good sense here.

Closing paragraph.

More discussion paragraphs

>>>>

Macroevolutionary studies frequently characterize the phylogenetic signal in phenotypes, and wish to compare the strength of that signal across traits. However, analytical tools for such comparisons have largely remained underdeveloped. In this study, we evaluated the efficacy of one commonly used parameter (Pagel's $\lambda$) to estimate the strength of phylogenetic signal in phenotypic traits, and evaluate the degree to which $\lambda$ correctly identifies known levels of phylogenetic signal. We find that the precision of $\lambda$ in estimating actual levels of phylogenetic signal is often inaccurate, and that biological interpretations of the strength of phylogenetic signal based on $\lambda$ are therefore compromised. We then propose a standardized effect size based on $Kappa$ ($Z_K$), which measures the strength of phylogenetic signal, and places it on a common scale for statistical comparison. Tests based on $Z_K$ provide a mechanism for formally comparing the strength of phylogenetic signal across datasets, in much the same manner as effect sizes may be used to summarize patterns in quantitative meta-analysis. Our approach extends the phylogenetic comparative toolkit to address hypotheses that compare the strength of phylogenetic signal various phenotypic traits, even when those traits are found in different evolutionary lineages or of in different units or scale.

@@

Evolutionary biologists commonly seek to describe the relative strength of phylogenetic signal in phenotypic traits, to determine the extent to which shared evolutionary history has influenced trait covariation among taxa. This is often accomplished by interpreting empirical estimates of $\lambda$; with smaller values signifying 'weak' phylogenetic signal, while larger values are interpreted as 'strong' phylogenetic signal [e.g., @DeMeester2019; @Pintanel2019;@Su2019]. Other approaches for interpreting $\lambda$ are more statistical. For instance, some have evaluated whether the observed $\lambda$ differs from some expected value through the use of confidence intervals [@Vandelook2019] or by performing likelihood ratio tests that compare the observed model fit to that obtained when $\lambda=0$ or $\lambda=1$ [@Freckleton_et_al2002; @Cooper2010; @Bose2019]. Additionally, qualitative comparisons of $\lambda$ estimates obtained from multiple phenotypic traits have been used to infer whether the strength of phylogenetic signal is greater in one trait as compared to another [e.g., @Liu2019; @Bai2019]. Indeed, statements regarding the strength of phylogenetic signal based on $\lambda$ are rather common in the evolutionary literature. For instance, of the 204 papers published in 2019 that estimated and reported Pagel's $\lambda$ (found in Google.scholar), 40% interpreted the strength of phylogenetic signal for at least one phenotypic trait. Further, because nearly half of the 1,572 $\lambda$ values reported were near 0 or 1 (Figure 1) where the biological interpretation of $\lambda$ is known, this percentage is even higher.

@@

It seems intuitive to interpret the strength of phylogenetic signal based on the value of $\lambda$, as $\lambda$ is a parameter on a bounded scale ($0\to1$) for which interpretation of its extremal points are understood. Specifically, $\lambda=0$ represents no phylogenetic signal, while $\lambda=1$ is phylogenetic signal as expected under Brownian motion. However, equating values of $\lambda$ directly to the strength of phylogenetic signal presumes two important statistical properties that have not been fully explored. First, it presumes that values of $\lambda$ can be precisely estimated, as biological inferences regarding the strength of phylogenetic signal depend on high accuracy in its estimation. Therefore, understanding the precision in estimating $\lambda$ is paramount. One study [@Boettiger_et_al2012] found that estimates of Pagel's $\lambda$ displayed less variation (i.e., greater precision) when data were simulated on a large phylogeny ($N=281$) as compared to a small one ($N=13$). From this observation it was concluded that insufficient data (i.e., the number of species) was the underlying cause of the increased variation across parameter estimates [@Boettiger_et_al2012]. Indeed, such a pattern is common with statistical estimators, as summary statistics and parameters are often more precise at greater sample sizes [@Cohen1988]. However, this conclusion also implies that the precision of $\lambda$ remains constant across its range ($\lambda = 0 \to 1$); an assumption that to date, has not been verified. Thus, despite widespread use of Pagel's [-@Pagel1999] $\lambda$ in macroevolutionary studies, at present, we lack a general understanding of the precision with which $\lambda$ can estimate levels of phylogenetic signal in phenotypic datasets. \hfill\break

Second, while estimates of $\lambda$ are within a bounded scale ($0\to1$), this does not *de-facto* imply that the estimated values of this parameter correspond to the actual strength of the underlying input signal in the data. For this to be the case, $\lambda$ must be a statistical effect size. Effect sizes are a measure the magnitude of a statistical effect in data, represented on a common scale [@Glass1976; @Cohen1988]. Effect sizes have widespread use in many areas of the quantiative sciences, as they represent measures that may be readily summarized across datasets as in meta-analyses [@HedgesOlkin1985; @Glass1976; @Arnqvist1995], or compared among datasets [e.g., @AdamsCollyer2016; @AdamsCollyer2019b]. Unfortunatley, not all model parameters and test statistics are effect sizes, and thus many summary measures must first be converted to standardized units (i.e., an effect size) for meaningful comparison [see @Rosenthal1994]. As a consequence, it follows that only if $\lambda$ is a statistical effect size can comparisons of estimates across datasets be interpretable. For the case of $\lambda$, this has not yet been explored.  \hfill\break

In this study, we evaluate the precision of Pagel's $\lambda$ for estimating known levels of phylogenetic signal in phenotypic data. We use computer simulations with differing numbers of species, differently shaped phylogenies, and differing input levels of phylogenetic signal, to explore the degree to which $\lambda$ correctly identifies known levels of phylogenetic signal, and under what circumstances. We find that estimates of $\lambda$ vary widely for a given input value of phylogenetic signal, and that the precision in estimating $\lambda$ is not constant across its range. Rather, there is decreased precision when input levels of phylogenetic signal are of intermediate strength. Additionally, the same estimated values of $\lambda$ may be obtained from datasets containing vastly different input levels of phylogenetic signal. Thus, $\lambda$ is not a reliable indicator of the strength of phylogenetic signal in phenotypic data. We then describe a standardized effect size for measuring the strength of phylogenetic signal in phenotypic datasets, and apply the concept to two common measures of phylogenetic signal: $\lambda$ and $Kappa$. Through simulations we find that the precision of effect sizes based on $\lambda$ ($Z_{\lambda}$) are less reliable than that those based on $Kappa$ ($Z_K$), implying that $Z_K$ is a more robust effect size measure. We also propose a two-sample test statistic that may be used to compare the strength of phylogenetic signal among datasets, and provide an empirical example to demonstrate its use. We conclude that estimates of phylogenetic signal using Pagel's $\lambda$ are often inaccurate, and thus interpreting strength of phylogenetic signal in phenotypic datasets based on this measure is compromised. By contrast, effect sizes obtained from $Kappa$ hold promise for characterizing phylogenetic signal, and for comparing the strength of phylogenetic signal across datasets.

@@

We conducted a series of computer simulations to evaluate the precision of Pagel's $\lambda$. Our primary simulations were based on pure-birth phylogenies; however, we also evaluated patterns on both balanced and pectinate trees to determine whether tree shape affected our findings (see Supporting Information). First we generated 50 pure-birth phylogenies at each of six different tree sizes, ranging from 32 to 1024 taxa ($n=2^5 - 2^{10}$). Next, we rescaled the simulated phylogenies by multiplying the internal branches by $\lambda_{in}$, using 21 intervals of 0.05 units across its range ($\lambda_{in} = 0.0 \to 1.0$), resulting in 1050 scaled phylogenies at each level of species richness ($n$). Continuous traits were then simulated on each phylogeny under a Brownian motion model of evolution to obtain datasets with differing levels of phylogenetic signal, that ranged from no phylogenetic signal (when $\lambda_{in} =0$), to phylogenetic signal reflecting Brownian motion (when $\lambda_{in} =1$). For each dataset we then estimated phylogenetic signal ($\lambda_{est}$), and calculated the variance of $\lambda$ ($\sigma^2_\lambda$) across datasets at each input level of phylogenetic signal and level of species richness as an estimate of precision. \hfill\break

We also evaluated the precision of $\lambda$ when estimated in PGLS regression and ANOVA (i.e., $Y\sim{X}$). Here, an independent variable $X$ was simulated on each rescaled phylogeny under a Brownian motion model of evolution (for PGLS regression). For phylogenetic ANOVA, random groups ($X$) were obtained by simulating a discrete (binary) character on each phylogeny. Next, the dependent variable was simulated in such a manner as to contain a known relationship with $X$ plus random error containing phylogenetic signal. This was accomplished as: $Y=\beta{X}+\epsilon$. Here, the association between $Y$ and $X$  was modeled using a range of values: $\beta=(0.0,0.25, 0.5, 0.75,1.0)$, and the residual error was modeled to contain phylogenetic signal simulated under a Brownian motion model of evolution on each rescaled phylogeny: $\epsilon=\mathcal{N}(\mu=0,\sigma=\sigma^2\mathbf{C})$: [see @Revell2010 for a similar simulation design]. The fit of the phylogenetic regression was estimated using maximum likelihood, and parameter estimates ($\beta_{est}$ and $\lambda_{est}$) were obtained. We then calculated precision estimates ($\sigma^2_\lambda$) at each input level of phylogenetic signal and level of species richness. \hfill\break

@@

*Results.* We found that the precision of $\lambda_{est}$ varied widely across simulation conditions. Predictably, precision improved as the number of species increased (Figure 2). This confirmed earlier findings of Boettiger et al. [-@Boettiger_et_al2012], and adhered to parametric statistical theory. However, in many cases the set of $\lambda_{est}$ spanned nearly the entire range of possible values (e.g., $n=32$; $\lambda_{in}=0.5$: $\lambda_{est}= 0.0\to 0.985$), revealing that estimates of $\lambda$ were not a reliable indicator of input phylogenetic signal. Importantly, the precision of $\lambda_{est}$ was not uniform across all levels of phylogenetic signal. The worst precision was observed at intermediate levels of phylogenetic signal ($\lambda_{in}\approx0.5$), while precision improved as input levels approached the extremes of $\lambda$'s range (i.e., $\lambda_{in}\rightarrow0$ & $\lambda_{in}\rightarrow1$). Thus, estimates of $\lambda$ were least reflective of the true input signal at intermediate values. Additionally, even at large levels of species richness, we found that the range of $\lambda_{est}$ still encompassed a substantial portion of possible values (e.g., $n=512$; $\lambda_{in}=0.5$: $\lambda_{est} = 0.32\to 0.68$). Likewise, the same $\lambda_{est}$ could be obtained from datasets containing vastly different input levels of phylogenetic signal (e.g., $n=512$; $\lambda_{est} = 0.5$; $\lambda_{in} = 0.25 \to 0.65$). These findings were particularly unsettling when considered in light of our literature survey. Over one quarter of the $\lambda$ estimates obtained in empirical studies (421 of 1,572) were between $\lambda=0.25$ and $\lambda=0.75$ (Figure 1). This range reflected the region that our simulations identified as being the least reliable in terms of accurately characterizing levels of phylogenetic signal, yet 30% of these mid-range empirical estimates were explicitly interpreted in terms of the strength of phylogenetic signal that they represented (i.e., weak, intermediate, strong phylogenetic signal). \hfill\break

Finally, when $\lambda$ was co-estimated with regression parameters in PGLS regression and ANOVA, the results of our simulations were quite similar. Here, regression parameters ($\beta$) were accurately estimated, confirming earlier findings of Revell [-@Revell2010] (see Supporting Information). However, estimates of phylogenetic signal ($\lambda$) were less precise (Figure 3; see also Supporting Information), and the spread of $\lambda_{est}$ was similar to that observed when $\lambda$ was estimated for only the dependent variable, as in Figure 2. Taken together, these findings reveal that $\lambda_{est}$ does not precisely characterize observed levels of phylogenetic signal in phenotypic datasets, and that biological interpretations of the strength of phylogenetic signal based on $\lambda$ may be highly inaccurate. \hfill\break

@@

Adams and Collyer [-@AdamsCollyer2019] suggested that the strength of phylogenetic signal could be represented as an effect size based on the $Kappa$ statistic and its empirical sampling distribution from permutation. Here we formalize that suggestion, resulting in an effect size of:

\begin{align}
    Z_{K}=\frac{log(K_{obs})-\hat\mu_{log(K)}}{\hat\sigma_{log(K)}}
\end{align}

where $K_{obs}$ is the observed phylogenetic signal, and $\hat\mu_K$ and $\hat\sigma_K$ are the mean and standard deviation of the empirical sampling distribution of $log(Kappa)$ obtained via permutation. Here the logarithm was used because $Kappa$ takes only positive values ($0\to\infty$) and its sampling distribution is log-normally distributed [for a similar transformation when calculating multivariate effect sizes see: Appendix 1 of @CollyerAdams2018]. Similarly, an effect size based on $\lambda$ could be envisioned as:

@@

In this case, $\lambda_{obs}$ and $\hat\sigma_\lambda$ are empirically derived using maximum likelihood, as permutation approaches have not been developed for evaluating $\lambda$. Note also that under the null hypothesis, no phylogenetic signal is expected [@Freckleton_et_al2002], and thus $E(\lambda)=0$ under this condition. \hfill\break

To evaluate the utility of $Z_K$ and $Z_\lambda$ we calculated both effect sizes for the simulated datasets generated above, and summarized the precision of each using its variance ($\sigma^2_{Z_K}$ and $\sigma^2_{Z_\lambda}$). Results are found in Figure 4 (additional results are found in the Supporting Information). Here two things are evident. First, estimates of $Z_K$ track the input phylogenetic signal in a more linear fashion than do estimates of $Z_\lambda$ (Figure 4A,B). Thus, actual changes in the strength of phylogenetic signal are reflected more evenly in the corresponding values of the effect size $Z_K$. Second, the precision of $Z_K$ is considerably more stable as compared with $Z_\lambda$. This may be seen by calculating the coefficients of variation for the set of precision estimates (i.e., $\sigma^2_{Z_K}$ and $\sigma^2_{Z_\lambda}$) across input levels of phylogenetic signal.  Here coefficients of variation in the precision of $Z_K$ were an order of magnitude smaller for than for $Z_\lambda$ (Figure 4C). This implied that estimates of the strength of phylogenetic signal were more reliable and robust when using $Z_K$ as compared with $Z_\lambda$. \hfill\break

@@

Once the magnitude of phylogenetic signal is characterized using $Z_K$, one may wish to compare such measures across datasets, to determine whether the strength of phylogenetic signal is significantly greater in one phenotypic trait as compared to another. As with other effect sizes derived from permutation distributions [e.g., @AdamsCollyer2016; @AdamsCollyer2019b], a two-sample test statistic may be calculated as:

@@

@@

# Discussion

It is common in comparative evolutionary studies to characterize the phylogenetic signal in phenotypic traits to determine the extent to which shared evolutionary history has generated trait covariation among taxa. However, while numerous analytical approaches may be used to quantify phylogenetic signal [e.g., @Abouheif1999; @Gittleman1990; @Blomberg_et_al2003; @Adams2014a; @Pagel1999], methods that explicitly measure the strength of phylogenetic signal, or facilitate comparisons among datasets, have remained underdeveloped. In this study, we evaluated the precision of one common measure, Pagel's $\lambda$, and explored its efficacy for characterizing the strength of phylogenetic signal in phenotypic data. Using computer simulations, we found that the precision of $\lambda$ increased with increasing sample sizes; a pattern noted previously [@Boettiger_et_al2012], and one that conformed with parametric statistical theory [@Cohen1988].  However, we also found that vastly different $\lambda$ estimates could be obtained from data containing the same level of phylogenetic signal, and that similar $\lambda$ estimates may be obtained from data containing differing levels of phylogenetic signal. Further, the precision of $\lambda$ varied with the strength of phylogenetic signal, where lower precision was observed when in data whose phylogenetic signal was of intemediate strength. From these findings we conclude that $\lambda$ is not a reliable indicator of the observed strength of phylogenetic signal in phenoytpic datasets, and that biological interpretations of the strength of signal based on this parameter may innacurately characterize such effects. \hfill\break

As an alternative, we described a standardized effect size ($Z$) for assessing the strength of phylogenetic signal. $Z$ expresses the magnitude of phylogenetic signal as a standard normal deviate, which is easily interpretable as the strength of phylogenetic signal relative to the mean. We applied this concept to both $\lambda$ and $Kappa$, and found that $Z_K$ was a better estimate of the strength of phylogenetic signal in phenotypic data. First, $Z_K$ was more precise than $Z_{\lambda}$, and variation in this precision was more consistent across the range of input levels of phylogenetic signal. Additionally, values of $Z_K$ more accurately tracked known levels of phylogenetic signal, with changes in the actual strength of phylogenetic signal reflected in a more linear fashion by concomite changes in the values of $Z_K$. Thus, $Z_K$ holds promise as a measure of the relative strength of phylogenetic signal that reflects the magnitude of this effect in phenotypic data. We therefore recommend that future studies interested in the strength of phylogenetic signal incorporate $Z_K$ as a statistical measure of this effect. \hfill\break

Based on the effect size $Z_K$, we then proposed a two-sample test, which provides a statistical means of determining whether the strength of phylogenetic signal is greater in one phenotypic trait as compared to another. Prior studies have summarized patterns of variation in phylogenetic signal across datasets using summary test values, such as $Kappa$ [e.g.,  @Blomberg_et_al2003]. However, $Kappa$ does not scale linearly with input levels of phylogenetic signal, and its variance increases (i.e., precision decreases) with increasing strength of phylogenetic signal [@Munkemuller_et_al2012; @DinizFilho2012: see also Supporting Information]. Thus, $Kappa$ should not be considered a standardized effect size that measures the strength of phylogenetic signal on a common scale. By contrast, converting $Kappa$ to $Z_K$ (via equation 2) alleviates these concerns, and facilitates formal statistical comparisons of the strength of signal across datasets. Thus when viewed from this perspective, the approach developed here is statistically aligned with other statistical approaches such as meta-analysis [sensu @HedgesOlkin1985; @Glass1976; @Arnqvist1995], where summary statistics across datasets are converted to standardized effect sizes for subsequent 'higher order' statistical summaries or comparisons. As such our approach enables evolutionary biologists to quantitatively examine the relative strength of phylogenetic signal across a wide range of phenotypic traits, and thus opens the door for future discoveries that inform on how phenotypic diversity accumulates in macroevolutionary time across the tree of life.  \hfill\break

One important advantage of the approach advocated here is that the resulting effect sizes ($Z_K$) are dimensionless, as the units of measurement cancel out during the calculation of $Z$ [@SokalRohlf2012]. Thus, $Z_K$ represents the strength of phylogenetic signal on a common and comparable scale -- measured in standard deviation units -- regardless of the initial units and original scale of the phenotypic variables under investigation. This means that the strength of phylogenetic signal may be compared across datsets for continuous phenotypic traits measured in different units and scale, because those units have been standardized through their conversion to $Z_K$. For example, our approach could be utilized to determine whether the strength of phylogenetic signal (say, in response to ecological differentiation) is stronger in morphological traits (linear traits: $mm$), physiological traits (metabolic rate: $\frac{O^2}{min}$), or behavioral traits (aggression: $\frac{\#{displays}}{second}$). Additionally, our method  is capable of comparing the strength of phylogenetic signal in traits of different dimensionality, as estimates of phylogenetic signal using $Kappa$ have been generalized for multivariate data [$K_{mult}$: see @Adams2014a]. Likewise, as $Z_K$ is a standard deviation unit (which accounts for variation, sample size, and the phylogeny), tests based on $\hat{Z}_{12}$ may be utilized for comparing the strength of phylogenetic signal across datasets containing a different number of species, and even for phenotypes obtained from species in different lineages, because their phylogenetic non-independence, and observed variation, are taken into account in the generation of the empirical sampling distribution via permutation.    \hfill\break

** Finally... Need Closing paragraph. **
